---
title: "Marketing Attribution"
output: html_notebook
author: Colin Priest
---

## load the R packages
```{r}
library(httr)
library(ggplot2)
library(scales)
library(datarobot)
library(ggthemes)
library(e1071)
library(plyr)
library(ggplot2) 
```


## Load the data from the historical marketing campaigns
```{r}
campaignData = read.csv('https://s3-ap-southeast-1.amazonaws.com/datarobotfiles/marketingDemo.csv', stringsAsFactors = FALSE)
head(campaignData)
```

## Connect to DataRobot and log in
```{r}
# get the login details
login = read.csv("c:/users/colin/Dropbox/login.csv", stringsAsFactors = FALSE)
username = login$Username[1]
apiToken = login$apiToken[1]
drKey = login$drKey[1]
rm(login)

# set the parameters
server = "https://app.datarobot.com"
predictionServer = "https://cfds.orm.datarobot.com"
# check that your account has API Access switched on
# connect to the DataRobot server
ConnectToDataRobot(endpoint = paste(server, "/api/v2", sep=""), token = apiToken)
```

## Build a DataRobot project that analyses the marketing campaign history
```{r}
target = "purchasePersonalLoan"
targetValue = "yes"
projectName = "Marketing Attribution"
autoPilotMetric = "AUC"
numWorkers = 20

# create the project
prj = SetupProject(campaignData, projectName, maxWait = 9999)
# start autopilot
SetTarget(project = prj, target = target, metric = autoPilotMetric, positiveClass = targetValue, 
          maxWait = 9999, mode = 'quick')
# use more (or less) workers
UpdateProject(prj$projectId, newProjectName = NULL, workerCount = numWorkers)
# wait for autopilot to finish
WaitForAutopilot(prj, verbosity = 0)

# pick a blueprint to use
allBlueprints = as.data.frame(GetAllModels(prj))
modelID = allBlueprints$modelId[which.max(allBlueprints$validationMetric)]
#
cat("Project model building completed\n")
```

## Score the training data and add the campaign costs
```{r}
# define which columns are the marketing touchpoints
touchPoints = c('facebook', 'googleAds', 'sms')

projectID = prj$projectId

# the url of the prediction server
url = paste0(predictionServer, '/api/v1/', projectID, '/', modelID, '/predict')
# send the data to the prediction server and get the predictions
myRequest = httr::POST(url,
  	                     add_headers("datarobot-key" = drKey),
    	                   c(authenticate(username, apiToken)),
      	                 body = rjson::toJSON(unname(split(campaignData, 1:nrow(campaignData)))), 
        	               httr::content_type_json())
response = content(myRequest, as = 'text', encoding = 'UTF-8')
response2 = jsonlite::fromJSON(response)
if (response2$status != '') error (response2)
predictions = jsonlite::fromJSON(response)$predictions
campaignOpt = campaignData[, c('CustomerID', target, touchPoints)]
campaignOpt$score = predictions$class_probabilities[,1]

campaignOpt$cost_facebook = 1 * (campaignOpt$facebook == 'yes')
campaignOpt$cost_googleAds = 0.1 * (campaignOpt$googleAds == 'yes')
campaignOpt$cost_sms = 0.085 * (campaignOpt$sms == 'yes')
campaignOpt$total_cost = campaignOpt$cost_facebook + campaignOpt$cost_googleAds + campaignOpt$cost_sms
campaignOpt$salesMargin = 2 * campaignOpt$score
campaignOpt$profit = campaignOpt$salesMargin - campaignOpt$total_cost

head(campaignOpt, 20)

```

## Plot the optimal campaign size
```{r}
campaignOpt = campaignOpt[order(campaignOpt$profit, decreasing = TRUE),]
campaignOpt$cumulativeProfit = cumsum(campaignOpt$profit)
campaignOpt$cumulativeSales = cumsum(campaignOpt$score)
campaignOpt$numberOfContacts = seq_len(nrow(campaignOpt))
campaignOpt$optimalCampaignSize = which.max(campaignOpt$cumulativeProfit)

g1 = ggplot(campaignOpt) + 
  theme_igray() +
  geom_line(aes(numberOfContacts, cumulativeProfit, colour = 'Profit')) +
  geom_point(aes(optimalCampaignSize, max(cumulativeProfit), colour = 'Optimal Campaign Size')) +
  xlab('Campaign Size (number of customers contacted)') +
  ylab('Total Profit') + ggtitle('Campaign Optimisation') +
  scale_x_continuous(labels = comma) + scale_y_continuous(labels = dollar) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(name="Legend",
    values=c(Profit='black', 'Optimal Campaign Size'='red'))
g1
```

## Build the what-if dataset for scoring
```{r}
newData = campaignData[campaignData$purchasePersonalLoan == 'yes',]

# generate all of the what-if possibilities
combinations = data.frame(bincombinations(length(touchPoints)))
names(combinations) = touchPoints
combinations$key = 'aaa'
combinations2 = combinations
combined = merge(combinations, combinations2, by = 'key')
rm(combinations, combinations2)
for (touchPoint in touchPoints)
{
  col1 = paste0(touchPoint, '.x')
  col2 = paste0(touchPoint, '.y')
  combined = combined[combined[, col2] <= combined[, col1],]
  combined[, touchPoint] = ifelse(combined[, col2] == 0, 'no', 'yes')
}
combined$key = unlist(lapply(seq_len(nrow(combined)), function(x) paste(combined[x,paste0(touchPoints, '.x')], collapse = '')))
combined = combined[, -(1 + seq_len(2 * length(touchPoints)))]

# generate the what-ifs
whatIfs = newData
whatIfs$key = unlist(lapply(seq_len(nrow(whatIfs)), function(x) paste(ifelse(whatIfs[x, touchPoints] == 'yes', 1, 0), collapse = '')))
whatIfs[, touchPoints] = NULL
whatIfs = merge(whatIfs, combined, by = 'key')
whatIfs = whatIfs[order(whatIfs$CustomerID, whatIfs$key),]
whatIfs$key = NULL
row.names(whatIfs) = NULL

head(whatIfs, 20)
```

## score the what-if scenarios
```{r}
projectID = prj$projectId

# the url of the prediction server
url = paste0(predictionServer, '/api/v1/', projectID, '/', modelID, '/predict')
# send the data to the prediction server and get the predictions
myRequest = httr::POST(url,
  	                     add_headers("datarobot-key" = drKey),
    	                   c(authenticate(username, apiToken)),
      	                 body = rjson::toJSON(unname(split(whatIfs, 1:nrow(whatIfs)))), 
        	               httr::content_type_json())
response = content(myRequest, as = 'text', encoding = 'UTF-8')
response2 = jsonlite::fromJSON(response)
if (response2$status != '') error (response2)
predictions = jsonlite::fromJSON(response)$predictions
whatIfs$score = predictions$class_probabilities[,1]

head(whatIfs, 20)
```

## calculate the Shapely value for each customer who purchased a product
```{r}
# only calculate the marketing contribution when a sales occurred
whatIfsYes = whatIfs[whatIfs$purchasePersonalLoan == 'yes',]
campaignDataYes = campaignData[campaignData$purchasePersonalLoan == 'yes',]

result = campaignDataYes
result = result[order(result$CustomerID),]

# get the baseline contribution - when all touchpoints are 'no'
temp = whatIfsYes
for (touchPoint in touchPoints)
	temp = temp[temp[, touchPoint] == 'no',]
result$contrib_baseline = temp$score

# calculate the Shapely value for each touch point
for (touchPoint in touchPoints)
{
  temp = whatIfsYes[, c('CustomerID', touchPoint, 'score')]
  n = names(temp)
  n[2] = 'touchPoint'
  n[3] = 'whatIfContribution'
  names(temp) = n
  temp2 = ddply(temp, .(CustomerID), summarize, n = length(whatIfContribution), cccxxx = sum(whatIfContribution[touchPoint == 'yes']) - sum(whatIfContribution[touchPoint == 'no']))
  temp2$cccxxx[temp2$n < 2] = 0
  col1 = paste0('contrib_', touchPoint)
  result[, col1] = temp2$cccxxx
  result[result[, touchPoint] == 'no', col1] = 0
  result[result$purchasePersonalLoan == 'no', col1] = 0
}

# rescale the Shapely value so that each row contribution adds to one sale
result$temp_total = result$contrib_baseline
for (touchPoint in touchPoints)
	result$temp_total = result$temp_total + result[, paste0('contrib_', touchPoint)]
result$contrib_baseline = result$contrib_baseline / result$temp_total
for (touchPoint in touchPoints)
	result[, paste0('contrib_', touchPoint)] = result[, paste0('contrib_', touchPoint)] / result$temp_total
result$temp_total = NULL
row.names(result) = NULL

head(result[, c('CustomerID', 'contrib_baseline', paste0('contrib_', touchPoints))], 20)
```

## summarise the Shapely values so that we can see the overall contribution of each marketing touchpoint
```{r}
df <- data.frame(
  touchpoint = c('None', touchPoints),
  value = c(sum(result$contrib_baseline), unlist(lapply(touchPoints, function(x) sum(result[, paste0('contrib_', x)]))))
  )
head(df)
df$touchpoint = factor(df$touchpoint, df$touchpoint)
df$percentage = df$value / sum(df$value) * 100.0

pie1 = ggplot(df, aes(x="", y=percentage, fill=touchpoint)) + geom_bar(width = 1, stat = "identity") +
   scale_fill_brewer(palette="RdBu") + coord_polar("y", start=0) + ggtitle('Marketing Attribution - Percentage') +
   xlab('') + ylab('') + labs(fill = 'Touch Point')
pie1
pie2 = ggplot(df, aes(x="", y=value, fill=touchpoint)) + geom_bar(width = 1, stat = "identity") +
   scale_fill_brewer(palette="RdBu") + coord_polar("y", start=0) + ggtitle('Marketing Attribution - Sales Count') +
   xlab('') + ylab('') + labs(fill = 'Touch Point')
pie2
```

